literature coding explore

### 2024-06-01 reports/lit_codes_explore.qmd
added code relations figures

code_freq <- read.csv(here("data/code_freq_032224.csv"))
codes_vars <- read.csv(here("data/code_w_vars_032224.csv"))
keywords <- read.csv(here("data/keywords_032224.csv"))
occur_all <- read.csv(here("data/code_configs_all_032224.csv"))
occur_br <- read.csv(here("data/code_configs_bridges_032224.csv"))
occur_ba <- read.csv(here("data/code_configs_barriers_032224.csv"))

@@

### Code appearance frequencies within code categories
Here, I have plotted the same information as the previous section, but as a percentage of the overall bridges and barriers groupings. For example, data bridges make up 37% of all bridge codes. The same pattern holds for the other sub-panels. For example, on the ecological page, abiotic indicators account for 22% of all ecological bridge codes.

#### Broader categories
```{r}
#| echo: false
#| message: false
#| warning: false
#| fig-width: 10
#| fig-height: 7
#| fig-align: center

@@

### Codes vs. study information

@@

::: panel-tabset
#### All codes

#### Bridges codes

#### Barriers codes

:::

### Keywords and literature type

::: panel-tabset
#### Keywords

#### Literature type

:::

>>>>

code_freq <- read.csv(here("data/code_freq_032224.csv"))
codes_vars <- read.csv(here("data/code_w_vars_032224.csv"))
keywords <- read.csv(here("data/keywords_032224.csv"))
occur_all <- read.csv(here("data/code_configs_all_032224.csv"))
occur_br <- read.csv(here("data/code_configs_bridges_032224.csv"))
occur_ba <- read.csv(here("data/code_configs_barriers_032224.csv"))
code_rel <- read.csv(here("data/code_rel.csv"))

@@

### Code appearance frequencies within code categories

Here, I have plotted the same information as the previous section, but as a percentage of the overall bridges and barriers groupings. For example, data bridges make up 37% of all bridge codes. The same pattern holds for the other sub-panels. For example, on the ecological page, abiotic indicators account for 22% of all ecological bridge codes.

::: panel-tabset
#### Broader categories

```{r}
#| echo: false
#| message: false
#| warning: false
#| fig-width: 10
#| fig-height: 7
#| fig-align: center

@@

@@

#### All code relations
This is a heatmap of all code relationships within articles. The number of co-ocurrences represents the number of times two codes co-occurered together in the same article.

```{r}
#| echo: false
#| message: false
#| warning: false
#| fig-width: 15
#| fig-height: 12
#| fig-align: center

code_rel_melt <- reshape::melt(code_rel)
code_rel_melt$variable <- gsub(".", " ", code_rel_melt$variable, fixed = TRUE)

code_rel_melt <- code_rel_melt %>%
  separate(Code.System, c("key", "code1"), "> ", convert = TRUE)

code_rel_melt$key <- str_trim(code_rel_melt$key)

colnames(code_rel_melt) <- c("brba_code1", "code1", "code2", "total")

ggplot(code_rel_melt, aes(reorder(code1, -total), reorder(code2, total))) +
geom_tile(aes(fill = total)) + 
scale_fill_gradient2(high = "red")+
ylab("")+
xlab("")+
labs(fill = "Number of co-occurences")+
theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 14))+
geom_text(aes(label = total))
```

#### Most frequent code relations
This heatmap represents the same data set as the last figure, but only includes co-occurrence values greater than or equal to 6, which is the 75% quantile of the co-occurrence totals. I have also added boxes around totals greater than 11, which was the 90% quantile of the co-occurrence total data set. Blue text represents data bridges and barriers, grey text represents ecological bridges and barriers, and orange text represents social bridges and barriers.

```{r}
#| echo: false
#| message: false
#| warning: false
#| fig-width: 12
#| fig-height: 10
#| fig-align: center

#add brba column to second code col
code_rel_melt2 <- code_rel_melt
code_rel_melt2$brba_code2 <- NA

code_rel_melt2$code2[code_rel_melt2$code2 == "Fine scale management appropriate for target species"] <- "Fine-scale management appropriate for target species"
code_rel_melt2$code2[code_rel_melt2$code2 == "Pre existing regulation and management institutions"] <- "Pre-existing regulation and management institutions"
code_rel_melt2$code2[code_rel_melt2$code2 == "Co management and communication"] <- "Co-management and communication"

for(i in 1:nrow(code_rel_melt2)){
  code_temp = code_rel_melt2$code2[i]

  brba_temp = code_rel_melt2 %>% filter(code1 == code_temp & code2 == code_temp)
  brba_code_temp = brba_temp$brba_code1[1]

  code_rel_melt2$brba_code2[i] <- brba_code_temp
}

code_rel_melt2$brba_code2[code_rel_melt2$code2 == "Target species is a prey specialist"] <- "Ecological bridge"

group_colors = c("Data Bridge" = "dodgerblue4", "Data Barrier" = "dodgerblue4", "Ecological Bridge" = "honeydew4", "Ecological Barrier" = "honeydew4", "Social Bridge" = "darkorange4", "Social Barrier" = "darkorange4")

code_rel_melt_h <- code_rel_melt2 %>%
  filter(total >= 6) %>%
  mutate(bor_box = ifelse(total >= 11, "TRUE", "FALSE"), 
        bor_box = as.factor(bor_box), 
        code1 = as.factor(code1), 
        code2 = as.factor(code2))

code_rel_melt_h$bor_box <- factor(code_rel_melt_h$bor_box, levels = c("TRUE", "FALSE"))

code_rel_melt_h <- code_rel_melt_h %>%
  mutate(code1 = reorder(code1, -total, FUN = mean),
         code2 = reorder(code2, total, FUN = mean))

unique_x_groups <- code_rel_melt_h %>%
distinct(code1, brba_code1) %>%
arrange(code1)

unique_y_groups <- code_rel_melt_h %>%
distinct(code2, brba_code2) %>%
arrange(code2)

x_colors <- setNames(group_colors[unique_x_groups$brba_code1], unique_x_groups$code1)
y_colors <- setNames(group_colors[unique_y_groups$brba_code2], unique_y_groups$code2)

ggplot(code_rel_melt_h, aes(reorder(code1, -total), reorder(code2, total))) +
geom_tile(aes(fill = total, colour = bor_box)) + 
scale_fill_gradient2(high = "red")+
ylab("")+
xlab("")+
labs(fill = "Number of co-occurences")+
theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 14, face = "bold", color = x_colors[levels(code_rel_melt_h$code1)]), 
      axis.text.y = element_text(size = 14, face = "bold", color = y_colors[levels(code_rel_melt_h$code2)]))+
geom_text(aes(label = total))+
scale_color_manual(values = c("black", "#00000000"))+
guides(color = "none")

#### Data code relations
This is a look at the data bridge and barrier code relations. Totals higher than or equal to 14 (90th quantile for data totals) are outlined in black.Here, red text represents codes that are barriers, and green text represents codes that are bridges.

```{r}
#| echo: false
#| message: false
#| warning: false
#| fig-width: 12
#| fig-height: 10
#| fig-align: center

code_rel_d <- read.csv(here("data/code_rel_d.csv"))

code_rel_melt_d <- reshape::melt(code_rel_d)
code_rel_melt_d$variable <- gsub(".", " ", code_rel_melt_d$variable, fixed = TRUE)

code_rel_melt_d <- code_rel_melt_d %>%
  separate(Code.System, c("key", "code1"), "> ", convert = TRUE)

code_rel_melt_d$key <- str_trim(code_rel_melt_d$key)

colnames(code_rel_melt_d) <- c("brba_code1", "code1", "code2", "total")

code_rel_melt2_d <- code_rel_melt_d
code_rel_melt2_d$brba_code2 <- NA
code_rel_melt2_d$brba_key1 <- NA
code_rel_melt2_d$brba_key2 <- NA

for(i in 1:nrow(code_rel_melt2_d)){
  code_temp = code_rel_melt2_d$code2[i]

  brba_temp = code_rel_melt2_d %>% filter(code1 == code_temp & code2 == code_temp)
  brba_code_temp = brba_temp$brba_code1[1]

  code_rel_melt2_d$brba_code2[i] <- brba_code_temp

  if(grepl("Bridge", code_rel_melt2_d$brba_code1[i])){
      code_rel_melt2_d$brba_key1[i] = "Bridge"
  } else if(grepl("Barrier", code_rel_melt2_d$brba_code1[i])){
      code_rel_melt2_d$brba_key1[i] = "Barrier"
  }

  if(grepl("Bridge", code_rel_melt2_d$brba_code2[i])){
      code_rel_melt2_d$brba_key2[i] = "Bridge"
  } else if(grepl("Barrier", code_rel_melt2_d$brba_code2[i])){
      code_rel_melt2_d$brba_key2[i] = "Barrier"
  }
}

code_rel_melt2_d <- code_rel_melt2_d %>%
  mutate(bor_box = ifelse(total >= 14, "TRUE", "FALSE"), 
        bor_box = as.factor(bor_box), 
        code1 = as.factor(code1), 
        code2 = as.factor(code2))

code_rel_melt2_d$bor_box <- factor(code_rel_melt2_d$bor_box, levels = c("TRUE", "FALSE"))

code_rel_melt2_d <- code_rel_melt2_d %>%
  mutate(code1 = reorder(code1, -total, FUN = mean),
         code2 = reorder(code2, total, FUN = mean))

brba_colors <- c("Bridge" = "darkolivegreen", "Barrier" = "red4")

unique_x_groups <- code_rel_melt2_d %>%
distinct(code1, brba_key1) %>%
arrange(code1)

unique_y_groups <- code_rel_melt2_d %>%
distinct(code2, brba_key2) %>%
arrange(code2)

x_colors <- setNames(brba_colors[unique_x_groups$brba_key1], unique_x_groups$code1)
y_colors <- setNames(brba_colors[unique_y_groups$brba_key2], unique_y_groups$code2)

ggplot(code_rel_melt2_d, aes(reorder(code1, -total), reorder(code2, total))) +
geom_tile(aes(fill = total, colour = bor_box)) + 
scale_fill_gradient2(high = "red")+
ylab("")+
xlab("")+
labs(fill = "Number of co-occurences")+
theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 14, face = "bold", color = x_colors[levels(code_rel_melt2_d$code1)]), 
      axis.text.y = element_text(size = 14, face = "bold", color = y_colors[levels(code_rel_melt2_d$code2)]))+
geom_text(aes(label = total))+
scale_color_manual(values = c("black", "#00000000"))+
guides(color = "none")

#### Ecological code relations
This is a look at the ecological bridge and barrier code relations. Totals higher than or equal to 5 (90th quantile for ecological totals) are outlined in black.Here, red text represents codes that are barriers, and green text represents codes that are bridges.

```{r}
#| echo: false
#| message: false
#| warning: false
#| fig-width: 12
#| fig-height: 10
#| fig-align: center

code_rel_e <- read.csv(here("data/code_rel_e.csv"))

code_rel_melt_e <- reshape::melt(code_rel_e)
code_rel_melt_e$variable <- gsub(".", " ", code_rel_melt_e$variable, fixed = TRUE)

code_rel_melt_e <- code_rel_melt_e %>%
  separate(Code.System, c("key", "code1"), "> ", convert = TRUE)

code_rel_melt_e$key <- str_trim(code_rel_melt_e$key)

colnames(code_rel_melt_e) <- c("brba_code1", "code1", "code2", "total")

code_rel_melt2_e <- code_rel_melt_e
code_rel_melt2_e$brba_code2 <- NA
code_rel_melt2_e$brba_key1 <- NA
code_rel_melt2_e$brba_key2 <- NA

code_rel_melt2_e$code2[code_rel_melt2_e$code2 == "Fine scale management appropriate for target species"] <- "Fine-scale management appropriate for target species"
code_rel_melt2_e$brba_code2[code_rel_melt2_e$code2 == "Target species is a prey specialist"] <- "Ecological bridge"

for(i in 1:nrow(code_rel_melt2_e)){
  code_temp = code_rel_melt2_e$code2[i]

  brba_temp = code_rel_melt2_e %>% filter(code1 == code_temp & code2 == code_temp)
  brba_code_temp = brba_temp$brba_code1[1]

  code_rel_melt2_e$brba_code2[i] <- brba_code_temp

  if(grepl("Bridge", code_rel_melt2_e$brba_code1[i])){
      code_rel_melt2_e$brba_key1[i] = "Bridge"
  } else if(grepl("Barrier", code_rel_melt2_e$brba_code1[i])){
      code_rel_melt2_e$brba_key1[i] = "Barrier"
  }

  if(grepl("Bridge", code_rel_melt2_e$brba_code2[i])){
      code_rel_melt2_e$brba_key2[i] = "Bridge"
  } else if(grepl("Barrier", code_rel_melt2_e$brba_code2[i])){
      code_rel_melt2_e$brba_key2[i] = "Barrier"
  }
}
code_rel_melt2_e$brba_code2[code_rel_melt2_e$code2 == "Target species is a prey specialist"] <- "Ecological bridge"
code_rel_melt2_e$brba_key2[code_rel_melt2_e$code2 == "Target species is a prey specialist"] <- "Bridge"

code_rel_melt2_e <- code_rel_melt2_e %>%
  mutate(bor_box = ifelse(total >= 5, "TRUE", "FALSE"), 
        bor_box = as.factor(bor_box), 
        code1 = as.factor(code1), 
        code2 = as.factor(code2))

code_rel_melt2_e$bor_box <- factor(code_rel_melt2_e$bor_box, levels = c("TRUE", "FALSE"))

code_rel_melt2_e <- code_rel_melt2_e %>%
  mutate(code1 = reorder(code1, -total, FUN = mean),
         code2 = reorder(code2, total, FUN = mean))

unique_x_groups <- code_rel_melt2_e %>%
distinct(code1, brba_key1) %>%
arrange(code1)

unique_y_groups <- code_rel_melt2_e %>%
distinct(code2, brba_key2) %>%
arrange(code2)

x_colors <- setNames(brba_colors[unique_x_groups$brba_key1], unique_x_groups$code1)
y_colors <- setNames(brba_colors[unique_y_groups$brba_key2], unique_y_groups$code2)

ggplot(code_rel_melt2_e, aes(reorder(code1, -total), reorder(code2, total))) +
geom_tile(aes(fill = total, colour = bor_box)) + 
scale_fill_gradient2(high = "red")+
ylab("")+
xlab("")+
labs(fill = "Number of co-occurences")+
theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 14, face = "bold", color = x_colors[levels(code_rel_melt2_e$code1)]), 
      axis.text.y = element_text(size = 14, face = "bold", color = y_colors[levels(code_rel_melt2_e$code2)]))+
geom_text(aes(label = total))+
scale_color_manual(values = c("black", "#00000000"))+
guides(color = "none")

#### Social code relations
This is a look at the social bridge and barrier code relations. Totals higher than or equal to 5 (90th quantile for social totals) are outlined in black.Here, red text represents codes that are barriers, and green text represents codes that are bridges.

```{r}
#| echo: false
#| message: false
#| warning: false
#| fig-width: 12
#| fig-height: 10
#| fig-align: center

code_rel_s <- read.csv(here("data/code_rel_s.csv"))

code_rel_melt_s <- reshape::melt(code_rel_s)
code_rel_melt_s$variable <- gsub(".", " ", code_rel_melt_s$variable, fixed = TRUE)

code_rel_melt_s <- code_rel_melt_s %>%
  separate(Code.System, c("key", "code1"), "> ", convert = TRUE)

code_rel_melt_s$key <- str_trim(code_rel_melt_s$key)

colnames(code_rel_melt_s) <- c("brba_code1", "code1", "code2", "total")

code_rel_melt2_s <- code_rel_melt_s
code_rel_melt2_s$brba_code2 <- NA
code_rel_melt2_s$brba_key1 <- NA
code_rel_melt2_s$brba_key2 <- NA

code_rel_melt2_s$code2[code_rel_melt2_s$code2 == "Pre existing regulation and management institutions"] <- "Pre-existing regulation and management institutions"
code_rel_melt2_s$code2[code_rel_melt2_s$code2 == "Co management and communication"] <- "Co-management and communication"

for(i in 1:nrow(code_rel_melt2_s)){
  code_temp = code_rel_melt2_s$code2[i]

  brba_temp = code_rel_melt2_s %>% filter(code1 == code_temp & code2 == code_temp)
  brba_code_temp = brba_temp$brba_code1[1]

  code_rel_melt2_s$brba_code2[i] <- brba_code_temp

  if(grepl("Bridge", code_rel_melt2_s$brba_code1[i])){
      code_rel_melt2_s$brba_key1[i] = "Bridge"
  } else if(grepl("Barrier", code_rel_melt2_s$brba_code1[i])){
      code_rel_melt2_s$brba_key1[i] = "Barrier"
  }

  if(grepl("Bridge", code_rel_melt2_s$brba_code2[i])){
      code_rel_melt2_s$brba_key2[i] = "Bridge"
  } else if(grepl("Barrier", code_rel_melt2_s$brba_code2[i])){
      code_rel_melt2_s$brba_key2[i] = "Barrier"
  }
}

code_rel_melt2_s <- code_rel_melt2_s %>%
  mutate(bor_box = ifelse(total >= 5, "TRUE", "FALSE"), 
        bor_box = as.factor(bor_box), 
        code1 = as.factor(code1), 
        code2 = as.factor(code2))

code_rel_melt2_s$bor_box <- factor(code_rel_melt2_s$bor_box, levels = c("TRUE", "FALSE"))

code_rel_melt2_s <- code_rel_melt2_s %>%
  mutate(code1 = reorder(code1, -total, FUN = mean),
         code2 = reorder(code2, total, FUN = mean))

unique_x_groups <- code_rel_melt2_s %>%
distinct(code1, brba_key1) %>%
arrange(code1)

unique_y_groups <- code_rel_melt2_s %>%
distinct(code2, brba_key2) %>%
arrange(code2)

x_colors <- setNames(brba_colors[unique_x_groups$brba_key1], unique_x_groups$code1)
y_colors <- setNames(brba_colors[unique_y_groups$brba_key2], unique_y_groups$code2)

ggplot(code_rel_melt2_s, aes(reorder(code1, -total), reorder(code2, total))) +
geom_tile(aes(fill = total, colour = bor_box)) + 
scale_fill_gradient2(high = "red")+
ylab("")+
xlab("")+
labs(fill = "Number of co-occurences")+
theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 14, face = "bold", color = x_colors[levels(code_rel_melt2_s$code1)]), 
      axis.text.y = element_text(size = 14, face = "bold", color = y_colors[levels(code_rel_melt2_s$code2)]))+
geom_text(aes(label = total))+
scale_color_manual(values = c("black", "#00000000"))+
guides(color = "none")
